R for Clinical Study Reports and Submission

### 2021-10-12 submission-package.Rmd
proofread chapter 12-13

In this chapter, we will first give a high-level overview of what
assets in the eCTD submission package we should focus on when
submitting R code.

Secondly, we will discuss how to prepare the proprietary R packages
(if any), and make them part of the submission package.

Lastly, we will provide reusable templates for updating ADRG and ARM,
so that the reviewers receive proper instructions to reproduce the
analysis results.

@@

The demo project (R package) we will prepare for submission is
available on GitHub at
[elong0527/esubdemo](https://github.com/elong0527/esubdemo).
You can download or clone it:

@@

- ADaM data in `.xpt` format: created by SAS
- `define.xml`: created by Pinnacle 21
- ADRG (ADaM Reviewer's Guide)
  - "Macro Programs" section: provide R and R package versions with a snapshot date
  - Appendix: provide step-by-step instructions for reviewers to
    reproduce the running environment and rerun analyses
- ARM (Analysis Results Metadata): provide the links between TLFs and
  analysis programs in tables

@@

- `r0pkgs.txt`: contains all internally developed R packages
- Other `.txt` files: each contains R code for a specific analysis

@@

For dependencies available from CRAN or public Git repositories,
you can declare them directly using the regular `Imports` and `Suggests`
syntax or the
[remotes dependency syntax](https://remotes.r-lib.org/articles/dependencies.html).

@@

@@

After we converted the R packages and R Markdown files into the appropriate
formats and verified that they can be restored and executed correctly,
we should update the ADRG (ADaM Reviewer's Guide) to provide
guidelines on how to use them.

>>>>

In this chapter, we will first give a high-level overview of what
assets in the eCTD submission package we should focus on when
submitting R code.
Secondly, we will discuss how to prepare the proprietary R packages
(if any), and make them part of the submission package.
Lastly, we will provide reusable templates for updating ADaM Reviewer's Guide (ADRG) and Analysis Results Metadata (ARM),
so that the reviewers receive proper instructions to reproduce the
analysis results.

@@

The demo project (R package) we will prepare for submission is called [`esubdemo`](https://github.com/elong0527/esubdemo.git) available on GitHub.
You can download or clone it:

@@

- ADaM data in `.xpt` format: created by SAS.
- `define.xml`: created by Pinnacle 21.
- ADRG: 
  - "Macro Programs" section: provide R and R package versions with a snapshot date.
  - Appendix: provide step-by-step instructions for reviewers to reproduce the running environment and rerun analyses.
- ARM: provide the links between TLFs and analysis programs in tables.

@@

- `r0pkgs.txt`: contains all internally developed R packages.
- Other `.txt` files: each contains R code for a specific analysis.

@@

For dependencies available from CRAN or public Git repositories,
you can declare them directly using the regular `Imports` and `Suggests`
syntax or the
[remotes dependency syntax](https://remotes.r-lib.org/articles/dependencies.html) in the `DESCRIPTION` file.

@@

To prepare R packages for submission, one needs to pack the packages into a text files, and then verify if the file only contains ASSCII.
With packed packages, one can unpack and install them from the text file, too.

@@

After we converted the R packages and R Markdown files into the appropriate
formats and verified that they can be restored and executed correctly,
we should update the ADRG to provide guidelines on how to use them.
